---
title: "Exploring the Palmer Penguins Dataset with R"
author: "Linda Arthur"
date: 2026-09-24

---


## Introduction

In this post, I explore the `palmerpenguins` dataset using R. I will use
data manipulation techniques from `dplyr` to answer two questions:

1. Which penguin has the highest bill length?
2. Which island has the most penguins?

I will also create a nested dataset that groups the penguin data by island
while keeping the other penguin details within each group.

```{r} 
#loading data and library
library(tidyverse)
library(palmerpenguins)
```

## 1. Which penguin with the highest bill length
```{r}
pen_highest_bill_len <- penguins |>
    arrange(desc(bill_length_mm)) |>
    slice(1)

pen_highest_bill_len
```

From the analysis, the penguins with the highest bill length is Gentoo on the Biscoe island

## 2a. Which island has the most penguins?

```{r}
island_max_pen <- penguins |>
    count(island,sort=TRUE) 
island_max_pen 
```

From the analysis above, the Biscoe island is the one withe the highest penguins of 168, followed by Dream(124) and Torgersen (52)

## 2b. Visualisaing the Number of penguins  per island
```{r} 
  ggplot(island_max_pen , aes(x = island, y = n)) +
  geom_col() +
  labs(
    title = "Number of Penguins by Island",
    x = "Island",
    y = "Number of Penguins"
  )

```


 ## 3. Nested dataset of Penguis by species and island
```{r}
pen_nested <- penguins |>
    group_by(species,island) |>
    nest()

pen_nested
pen_nested$data[[1]]
```

I created a nested dataset that groups the penguins by species and island while keeping the other penguin information within each island as a list-column. The last line select the information about the first penguin in the dataset and its information which is Adelie  and on the Torgersen island.